---
title: "ccbr1045_trajectoryAnalysis"
author: "Da"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### y_x.rds_1.2_tabulaMuris2020.3.4.rds is the four sample merged seurat object with added tabula Muris cell identity annotation
```{r}
setwd("~/Desktop/active_projects/ccbr1045/") 
#original data source: biowulf /data/CCBR/projects/ccbr1045/scrna_guohao2/SingleR_testing/y_x.rds_1.2_tabulaMuris2020.3.4.rds
y_x.rds_1.2 = readRDS("~/Desktop/active_projects/seuratObject/y_x.rds_1.2_tabulaMuris2020.3.4.rds")
#original data source: biowulf /data/CCBR/projects/ccbr1045/scrna_guohao2/integration/merged/y-x/microgliaReshape.rds
microgliaReshape = readRDS("~/Desktop/active_projects/seuratObject/microgliaReshape.rds")

save.image(file = "monocle.RData")
```

#### subset samples and clusters
```{r}
microglia_Arf1 = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="Arf1_KO"&microgliaReshape$new_clusters%in%c(1,2,3,4,5,6,7,8,9,10,11,12)))



microglia_Arf1_main = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="Arf1_KO"&microgliaReshape$new_clusters%in%c(1,2,3,4,5)))

microglia_WT_main = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="WT"&microgliaReshape$new_clusters%in%c(1,2,3,4,5)))


microglia_Arf1_IFNg_main = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="Arf1_IFNg_KO"&microgliaReshape$new_clusters%in%c(1,2,3,4,5)))


microglia_IFNg_main = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="IFNg_KO"&microgliaReshape$new_clusters%in%c(1,2,3,4,5)))


microglia_Arf1_main2 = subset(x = microgliaReshape, cells = which(microgliaReshape$SampleName=="Arf1_KO"&microgliaReshape$new_clusters%in%c(1,2,4,5)))
microglia_Arf1_main2@meta.data
```

#### function for pre-processing, reduce dimensionality, clustering for monocle
```{r}
myTrajectory = function(so){
  data = as(as.matrix(so@assays$RNA@counts), 'sparseMatrix')
  pd <- new('AnnotatedDataFrame', data = so@meta.data)
  fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
  fd <- new('AnnotatedDataFrame', data = fData)
  my_cds <- newCellDataSet(data,
                              phenoData = pd,
                              featureData = fd,
                              lowerDetectionLimit = 0.5,
                              expressionFamily = negbinomial.size())
  my_cds <- estimateSizeFactors(my_cds)
  my_cds <- estimateDispersions(my_cds)
  my_cds <- detectGenes(my_cds, min_expr = 0.1)
  pData(my_cds)$UMI <- Matrix::colSums(exprs(my_cds))
  disp_table <- dispersionTable(my_cds)
  table(disp_table$mean_expression>=0.1)
  unsup_clustering_genes <- subset(disp_table, mean_expression >= 0.1)
  my_cds <- setOrderingFilter(my_cds, unsup_clustering_genes$gene_id)
  
  expressed_genes <- row.names(subset(fData(my_cds), num_cells_expressed >= 10))
  my_cds_subset <- my_cds[expressed_genes, ]
  
  my_cds_subset <- detectGenes(my_cds_subset, min_expr = 0.1)
  fData(my_cds_subset)$use_for_ordering <- fData(my_cds_subset)$num_cells_expressed > 0.05 * ncol(my_cds_subset)
  table(fData(my_cds_subset)$use_for_ordering)
  
  #plot_pc_variance_explained(my_cds_subset, return_all = FALSE) # takes long time

  my_cds_subset <- reduceDimension(my_cds_subset,max_components = 2,norm_method = 'log',num_dim = 10,
                                   reduction_method ='tSNE',verbose = TRUE)
  my_cds_subset <- clusterCells(my_cds_subset, verbose = FALSE)
  plot_rho_delta(my_cds_subset, rho_threshold = 2, delta_threshold = 10)
  my_cds_subset <- clusterCells(my_cds_subset,rho_threshold = 2,delta_threshold = 10,skip_rho_sigma = T,verbose = FALSE)
  
  clustering_DEG_genes <- differentialGeneTest(my_cds_subset,fullModelFormulaStr = '~Cluster',cores = 2)
  
  clustering_DEG_genes %>% arrange(qval) %>% head()
  my_ordering_genes <- row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:1000]
  my_cds_subset <- setOrderingFilter(my_cds_subset, ordering_genes = my_ordering_genes)
  my_cds_subset <- reduceDimension(my_cds_subset, method = 'DDRTree')
  my_cds_subset <- orderCells(my_cds_subset)
  
  plot_cell_trajectory(my_cds_subset, color_by = "new_clusters",cell_size = 1)
  plot_cell_trajectory(my_cds_subset, color_by = "State",cell_size = 1)
  
  return(my_cds_subset)

}
```

```{r}
my_cds_subset_Arf1 = myTrajectory(so = microglia_Arf1)

my_cds_subset_Arf1_main = myTrajectory(so = microglia_Arf1_main)

```

#### set the root state and order cells in pseudotime
```{r}
my_cds_subset_Arf1_main = myTrajectory(so = microglia_Arf1_main)

plot_cell_trajectory(my_cds_subset_Arf1, color_by = "new_clusters",cell_size = 0.6 )
plot_cell_trajectory(my_cds_subset_Arf1, color_by = "State",cell_size = 0.6)

plot_cell_trajectory(my_cds_subset_Arf1_main, color_by = "new_clusters",cell_size = 0.6)
plot_cell_trajectory(my_cds_subset_Arf1_main, color_by = "State",cell_size = 0.6)

my_cds_subset_Arf1_main <- orderCells(my_cds_subset_Arf1_main,root_state=7)

my_pseudotime_de_Arf1_main <- differentialGeneTest(my_cds_subset_Arf1_main,fullModelFormulaStr = "~sm.ns(Pseudotime)",cores = 2)

my_pseudotime_de_Arf1_main %>% arrange(qval) %>% head()
my_pseudotime_de_Arf1_main %>% arrange(qval) %>% head() %>% select(gene_short_name) -> my_pseudotime_gene_Arf1_main

plot_cell_trajectory(my_cds_subset_Arf1_main, color_by = "Pseudotime",cell_size = 0.6)

plot_cell_trajectory(my_cds_subset_Arf1_main, markers = "Apoe",use_color_gradient = T,cell_size = 0.6) +
scale_color_gradient2(low="navy", mid="white", high="red", 
                       midpoint=0)


```

#### set the root state and order cells in pseudotime

```{r}
my_cds_subset_WT_main = myTrajectory(so = microglia_WT_main)

plot_cell_trajectory(my_cds_subset_WT_main, color_by = "new_clusters",cell_size = 0.6)
plot_cell_trajectory(my_cds_subset_WT_main, color_by = "State",cell_size = 0.6)

my_cds_subset_WT_main <- orderCells(my_cds_subset_WT_main,root_state=3)

my_pseudotime_de_WT_main <- differentialGeneTest(my_cds_subset_WT_main,fullModelFormulaStr = "~sm.ns(Pseudotime)",cores = 2)

my_pseudotime_de_WT_main %>% arrange(qval) %>% head()
my_pseudotime_de_WT_main %>% arrange(qval) %>% head() %>% select(gene_short_name) -> my_pseudotime_gene_WT_main

plot_cell_trajectory(my_cds_subset_WT_main, color_by = "Pseudotime",cell_size = 0.6)
```

#### set the root state and order cells in pseudotime
```{r}
my_cds_subset_Arf1_main2b = myTrajectory(so = microglia_Arf1_main2)
plot_cell_trajectory(my_cds_subset_Arf1_main2b, color_by = "new_clusters",cell_size = 1)
plot_cell_trajectory(my_cds_subset_Arf1_main2b, color_by = "State",cell_size = 1)

my_cds_subset_Arf1_main2b <- orderCells(my_cds_subset_Arf1_main2b,root_state=2)

my_pseudotime_de_Arf1_main2b <- differentialGeneTest(my_cds_subset_Arf1_main2b,fullModelFormulaStr = "~sm.ns(Pseudotime)",cores = 2)

my_pseudotime_de_Arf1_main2b %>% arrange(qval) %>% head()
my_pseudotime_de_Arf1_main2b %>% arrange(qval) %>% head() %>% select(gene_short_name) -> my_pseudotime_gene_Arf1_main2b

plot_cell_trajectory(my_cds_subset_Arf1_main2b, color_by = "Pseudotime",cell_size = 1)

```


```{r}
my_cds_subset_Arf1_main2 = myTrajectory(so = microglia_Arf1_main2)
plot_cell_trajectory(my_cds_subset_Arf1_main2, color_by = "new_clusters",cell_size = 1)
plot_cell_trajectory(my_cds_subset_Arf1_main2, color_by = "State",cell_size = 1)

```


```{r}
data = as(as.matrix(microglia_Arf1_main@assays$RNA@counts), 'sparseMatrix')
pd <- new('AnnotatedDataFrame', data = microglia_Arf1_main@meta.data)

fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
fd <- new('AnnotatedDataFrame', data = fData)
my_cds <- newCellDataSet(data,
                              phenoData = pd,
                              featureData = fd,
                              lowerDetectionLimit = 0.5,
                              expressionFamily = negbinomial.size())

my_cds <- estimateSizeFactors(my_cds)
my_cds <- estimateDispersions(my_cds)
my_cds <- detectGenes(my_cds, min_expr = 0.1)
head(fData(my_cds))
head(pData(my_cds))

pData(my_cds)$UMI <- Matrix::colSums(exprs(my_cds))
disp_table <- dispersionTable(my_cds)
head(disp_table)

table(disp_table$mean_expression>=0.1)
unsup_clustering_genes <- subset(disp_table, mean_expression >= 0.1)
my_cds <- setOrderingFilter(my_cds, unsup_clustering_genes$gene_id)
plot_ordering_genes(my_cds)


expressed_genes <- row.names(subset(fData(my_cds), num_cells_expressed >= 10))
my_cds_subset <- my_cds[expressed_genes, ]
my_cds_subset
head(pData(my_cds_subset))


my_cds_subset <- detectGenes(my_cds_subset, min_expr = 0.1)
fData(my_cds_subset)$use_for_ordering <- fData(my_cds_subset)$num_cells_expressed > 0.05 * ncol(my_cds_subset)
table(fData(my_cds_subset)$use_for_ordering)

#plot_pc_variance_explained(my_cds_subset, return_all = FALSE) # takes long time

my_cds_subset <- reduceDimension(my_cds_subset,max_components = 2,norm_method = 'log',num_dim = 10,reduction_method = 'tSNE',verbose = TRUE)
my_cds_subset <- clusterCells(my_cds_subset, verbose = FALSE)
plot_rho_delta(my_cds_subset, rho_threshold = 2, delta_threshold = 10)
my_cds_subset <- clusterCells(my_cds_subset,rho_threshold = 2,delta_threshold = 10,skip_rho_sigma = T,verbose = FALSE)



table(pData(my_cds_subset)$Cluster)
plot_cell_clusters(my_cds_subset)
head(pData(my_cds_subset))


clustering_DEG_genes <- differentialGeneTest(my_cds_subset,fullModelFormulaStr = '~Cluster',cores = 12)
dim(clustering_DEG_genes)
```


The single-cell trajectory reconstructed by Monocle 2 for Arf1_KO cells. Cells start at the root and progress to one of three alternative outcomes, denoted by 1, 2, and 3. To reach these fates, cells must pass through branch point 1. Cells that do not proceed to F1 must then choose between F2 or F3 at branch point B2.

```{r}
#saveRDS(object = my_cds_subset, file = "~/Desktop/active_projects/ccbr1045/my_cds_subset1.rds")

clustering_DEG_genes %>% arrange(qval) %>% head()
my_ordering_genes <- row.names(clustering_DEG_genes)[order(clustering_DEG_genes$qval)][1:1000]
my_cds_subset <- setOrderingFilter(my_cds_subset, ordering_genes = my_ordering_genes)
my_cds_subset <- reduceDimension(my_cds_subset, method = 'DDRTree')
my_cds_subset <- orderCells(my_cds_subset)

#saveRDS(object = my_cds_subset, file = "~/Desktop/active_projects/ccbr1045/my_cds_subset2.rds")

#pseudotime trajectories calculated
#plot_cell_trajectory(my_cds_subset, color_by = "SampleName")
plot_cell_trajectory(my_cds_subset, color_by = "new_clusters",cell_size = 1)
#plot_cell_trajectory(my_cds_subset, color_by = "orig.ident")
head(pData(my_cds_subset))
```

```{r}
#plot_cell_trajectory(my_cds_subset, color_by = "State",cell_size = 1)+ facet_wrap(~State, nrow = 1)
plot_cell_trajectory(my_cds_subset, color_by = "State",cell_size = 1)
```
Since we don't have a timeseries, you need to set the root based on where certain marker genes are expressed, in this case, the homeostatic genes, cluster1 and cluster2, which corresponds to State 7. 
```{r}
#library(monocle)
my_cds_subset <- orderCells(my_cds_subset,root_state=7)

my_pseudotime_de <- differentialGeneTest(my_cds_subset,fullModelFormulaStr = "~sm.ns(Pseudotime)",cores = 2)

my_pseudotime_de %>% arrange(qval) %>% head()
my_pseudotime_de %>% arrange(qval) %>% head() %>% select(gene_short_name) -> my_pseudotime_gene

plot_cell_trajectory(my_cds_subset, color_by = "Pseudotime",cell_size = 1)
```




```{r}
#"A" stand for top 6 genes of affecting the fate decisions
A=c("Rpl37a","Cst3","Rps21","Rpl39","P2ry12","Rps20")
B=c("Apoe","Cst3","Tmem119","Spp1","P2ry12","Cx3cr1")
my_pseudotime_gene <-B
plot_genes_in_pseudotime(my_cds_subset[my_pseudotime_gene,], color_by = "new_clusters",cell_size = 0.3)

plot_cells(my_cds_subset,
           genes=A,
           label_cell_groups=FALSE,
           show_trajectory_graph=FALSE)

save.image(file = "monocle.RData")


```


```{r}
#Calculate the heat map of the top 30 genes
library(dplyr)
my_pseudotime_de %>% arrange(qval) %>% head(50) %>% select(gene_short_name) -> gene_to_cluster
gene_to_cluster <- gene_to_cluster$gene_short_name
my_pseudotime_cluster <- plot_pseudotime_heatmap(my_cds_subset[gene_to_cluster,],num_clusters = 5,cores = 2,show_rownames = TRUE,return_heatmap = TRUE)

```
